Testing atlantisom: true biomass output from Atlantis

### 2019-05-20 Atlantisom2SSworkflowtest.Rmd
try lf with 3 species, still too big

Note: on Sarah's laptop, CCA took ~40 minutes to return from `run_truth` and NOBA took an hour and 26 minutes to return. Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.

@@

```{r surveyBbased}

@@

```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
# plot some comparisons with Atlantis output

# read Atlantis output files
  atBtxt2 <- read.table(here(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

@@

#all species comparison, time intervals hardcoded for NEUS and NOBA
compareB <-ggplot() +
  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

```{r surveyNbased}

@@

@@

The next steps of sampling can get us a proper weight at age. Here again a census to generate length comps, which need an average weight at age and carry it forward. First we `sample_fish` (nums and weight components):

```{r biolsampling}

@@

# apply default sample fish as before to get numbers
numsallhigh <- sample_fish(survey_testNall, effNhigh)

# aggregate true resn per survey design
aggresnall <- aggregateDensityData(dat = truth$resn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

# aggregate true structn per survey design
aggstructnall <- aggregateDensityData(dat = truth$structn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)

resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)

@@

Then we calculate lengths with `calc_age2length`

@@

length_census <- calc_age2length(structn = structnall,
                                 resn = resnall,
                                 nums = numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

Now we can test the full `sample_survey_biomass` from nums to biomass index:

@@

wtage <- length_census$muweight %>%
  select(species, agecl, wtAtAge = atoutput)

surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)

@@

#all species comparison, time intervals hardcoded for NEUS and NOBA
compareB_N <-ggplot() +
  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

>>>>

Note: on Sarah's laptop, CCA took ~2.5 hours to return from `run_truth` and NOBA took an hour and 26 minutes to return. Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.

@@

```{r surveyBbased, eval=F}

@@

```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 36 species.", message=FALSE, warning=FALSE}
# plot some comparisons with Atlantis output

# read Atlantis output files
atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

surveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

@@

#all species comparison, time intervals hardcoded for 5 steps per year
compareB <-ggplot() +
  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

```{r surveyNbased, eval=F}

@@

#save for later use, takes a long time to generate
saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

@@

We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:

```{r truenatagecl, eval=FALSE}
# We need an effective N for the sample fish function, setting equal to actual N
# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

numsallhigh <- sample_fish(survey_testNall, effNhigh)

#save for later use, takes a long time to generate
saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))

The next steps of sampling can get us a proper weight at age. Here again a census to generate length comps, which need an average weight at age and carry it forward. First we `sample_fish` (nums and weight components).

On a full model run this takes far too long with all species, so we cut it down to a few assessment species here:

```{r ss-biolsampling}

# get only assessed species, hardcoded here for CCA sardine, hake, bocaccio
spp.name <- funct.group.names[funct.group.names %in% c("Pacific_sardine",
                                                       "Mesopel_M_Fish",
                                                       "Bocaccio_rockfish")]

survey_ssN <- create_survey(dat = truth$nums,
                                 time = timeall,
                                 species = spp.name,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

@@

# apply default sample fish as before to get numbers
numssshigh <- sample_fish(survey_ssN, effNhigh)

# aggregate true resn per survey design
aggresnss <- aggregateDensityData(dat = truth$resn,
                                 time = timeall,
                                 species = spp.name,
                                 boxes = boxall)

# aggregate true structn per survey design
aggstructnss <- aggregateDensityData(dat = truth$structn,
                                 time = timeall,
                                 species = spp.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
structnss <- sample_fish(aggstructnss, effNhigh, sample = FALSE)

resnss <-  sample_fish(aggresnss, effNhigh, sample = FALSE)

@@

Then we calculate lengths with `calc_age2length`. This will probably not work for the full run with all species outputs 5x per year, but should be ok with 3 species.

@@

length_census_ss <- calc_age2length(structn = structnss,
                                 resn = resnss,
                                 nums = numssshigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

```{r savelengths}
#save for later use, takes a long time to generate
saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
```

Now we can test the full `sample_survey_biomass` from nums to biomass index with these three species:

@@

length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))

wtage <- length_census_ss$muweight %>%
  select(species, agecl, wtAtAge = atoutput)

surveyB_fromN <- sample_survey_biomass(survey_ssN, surv_cv, wtage)

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% levels(surveyB_fromN$species))

#all species comparison, time intervals hardcoded for 5 steps per year
compareB_N <-ggplot() +
  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)
